Project Description -- Erosion, Transport, Anomalous Diffusion, and Collapse in Granular and Porous Media

### 2019-11-18 docs/proposals/NSF/ProjDesc/ProjDesc.tex
NM

%----------------------------------------------------------------------%
\subsection{Governing Equations}
At the finest scale, the fluid flow intervening between individual grains of a porous medium is governed by the incompressible Stokes equations \begin{equation}
\label{eqn:StokesEq}
\begin{split}
\mu \Delta \uu = \grad p,	&\hspace{20pt} \xx \in \Omega, \\
\grad \cdot \uu = 0, 		&\hspace{20pt} \xx \in \Omega,
\end{split}
\end{equation}
where $\uu$ is the velocity field, $p$ the pressure field, $\mu$ the
fluid viscosity, and $\Omega$ is the fluid domain. Linear elliptic PDEs,
such as the Stokes equations, can be reformulated as a boundary integral
equation (BIE).  BIEs have several numerical advantages including a
dimension reduction, automatic mass conservation, and high-order
accuracy. An equivalent BIE formulation of the Stokes equations in
Eq.~\eqref{eqn:StokesEq} is
\begin{align}
  \label{eqn:BIE}
  \ff(\xx) = -\frac{1}{2}\eeta(\xx) + \frac{1}{\pi} \int_{\bd\Omega}
    \frac{\rr \cdot \nn}{\rho^2} \frac{\rr \otimes \rr}{\rho^2}
    \eeta(\yy) ds_\yy, \quad \xx \in \bd\Omega,
\end{align}
where $\ff$ is a Dirichlet boundary condition, $\rr = \xx - \yy$, $\rho = \|\rr\|$, and $\nn$ is the outward unit normal of $\bd\Omega$.  Once~\eqref{eqn:BIE} is solved for the density function $\eeta$, important quantities such as the velocity, pressure, vorticity, and deformation tensor can be computed in the fluid bulk.  For example, the velocity at $\xx \in \Omega$ is 
\begin{align}
  \uu(\xx) = \frac{1}{\pi} \int_{\bd\Omega}
    \frac{\rr \cdot \nn}{\rho^2} \frac{\rr \otimes \rr}{\rho^2}
    \eeta(\yy) ds_\yy.
\end{align}

@@

\paragraph{Relevance to the Computational Mathematics Synopsis:}
\begin{itemize}[noitemsep]
  \item Computational plays an essential role to simulate erosion of porous medium.

  \item The proposed research is directly applicable to several geophysical applications.

  \item The proposed research is a collaborative effort and will include graduate, undergraduate, and K--12 students
\end{itemize}

@@

\paragraph{PIs' Qualifications:} PI Moore joined the Department of Mathematics at Florida State University (FSU) in 2014, and is also a Research Associate in FSU's Geophysical Fluid Dynamics Institute (GFDI).  His expertise includes modeling dynamic interactions between fluids and structures, especially in geophysical and biologically motivated settings. PI Quaife joined the Department of Scientific Computing at FSU in 2015 and is also a Research Associate in GFDI. His expertise includes high-fidelity and efficient numerical methods for fluid dynamic in complex geometries.

The PIs co-authored a paper describing numerical methods for simulating erosion~\cite{Quaife2018} in a viscous fluid. Very recently, a second paper has been submitted to the Journal of Fluid Mechanics~\cite{chi-moo-qua2019}, where additional numerical methods are developed and characterizations of transport are presented.

>>>>

%----------------------------------------------------------------------%
\subsection{Governing Equations}

Here, we first outline the theoretical foundation to be used in later sections of the proposal. At the finest scale, the fluid flow intervening between individual grains of a porous medium is governed by the incompressible Stokes equations \begin{equation}
\label{eqn:StokesEq}
\begin{split}
\mu \Delta \uu = \grad p,	&\hspace{20pt} \xx \in \Omega, \\
\grad \cdot \uu = 0, 		&\hspace{20pt} \xx \in \Omega,
\end{split}
\end{equation}
where $\uu$ is the velocity field, $p$ the pressure field, $\mu$ the
fluid viscosity, and $\Omega$ is the fluid domain. Linear elliptic PDEs,
such as the Stokes equations, can be reformulated as a boundary integral
equation (BIE).  BIEs have several numerical advantages including a
dimension reduction, automatic mass conservation, and high-order
accuracy. An equivalent BIE formulation of the Stokes equations in
Eq.~\eqref{eqn:StokesEq} is
\begin{align}
  \label{eqn:BIE}
  \ff(\xx) = -\frac{1}{2}\eeta(\xx) + \frac{1}{\pi} \int_{\bd\Omega}
    \frac{\rr \cdot \nn}{\rho^2} \frac{\rr \otimes \rr}{\rho^2}
    \eeta(\yy) ds_\yy, \quad \xx \in \bd\Omega,
\end{align}
where $\ff$ is a Dirichlet boundary condition, $\rr = \xx - \yy$, $\rho = \|\rr\|$, and $\nn$ is the outward unit normal of $\bd\Omega$.  Once~\eqref{eqn:BIE} is solved for the density function $\eeta$, important quantities such as the velocity, pressure, vorticity, and deformation tensor can be computed in the fluid bulk.  For example, the velocity at $\xx \in \Omega$ is 
\begin{align}
  \uu(\xx) = \frac{1}{\pi} \int_{\bd\Omega}
    \frac{\rr \cdot \nn}{\rho^2} \frac{\rr \otimes \rr}{\rho^2}
    \eeta(\yy) ds_\yy.
\end{align}

@@

\paragraph{Relevance to the Computational Mathematics Synopsis:}
\begin{itemize}[noitemsep]
  \item Efficient, high-fidelity computational methods form the core of the proposed work of porous-media erosion. Large ensembles of the high-fidelity simulations will inform the coarse-grained models that will be developed in the later stages of the proposed work.
  \item The proposed research is interdisciplinary by nature, with immediate applications in the fields of geophysics and environmental sciences. The PIs already have experience collaborating with scientists from other disciplines through FSU's Geophysical Fluid Dynamics Institute, an interdisciplinary unit comprised of a diverse set of faculty from home departments in example Earth, Atmospheric, and Ocean Sciences, Geology, Physics, Civil Engineering, Environmental Engineering, and others.
  \item The proposed research is a collaborative effort with a focus on training future generations of scientists and mathematicians. The proposal will include graduate, undergraduate, and K--12 students, with close cooperation between them.
\end{itemize}

@@

\paragraph{PIs' Qualifications:} PI Moore joined the Department of Mathematics at Florida State University (FSU) in 2014. He is also a Research Associate in FSU's Geophysical Fluid Dynamics Institute (GFDI).  His expertise includes modeling dynamic interactions between fluids and structures, especially in geophysical and biologically motivated settings. PI Quaife joined the Department of Scientific Computing at FSU in 2015 and is also a Research Associate in GFDI. His expertise includes high-fidelity and efficient numerical methods for fluid dynamic in complex geometries.

The PIs co-authored a paper in the {\em Journal of Computational Physics} describing numerical methods for simulating erosion~\cite{Quaife2018} in the Stokes flow regime, which is the regime relevant for the currently proposed applications of groundwater flow. Very recently the PIs, along with postdoctoral scholar Shang-Huan Chiu, submitted a second paper on newly developed methods for precise characterization of transport and dispersion within the special class of porous-media configurations formed by erosion \cite{chi-moo-qua2019}.

%Very recently, a second paper has been submitted to the Journal of Fluid Mechanics~\cite{chi-moo-qua2019}, where additional numerical methods are developed and characterizations of transport are presented.
